---
output: html_document
bibliography: references.bib
csl: elsevier-with-titles.csl
link-citations: true
---

```{r setup, include=FALSE}
library(knitcitations)
cite_options(citation_format = 'pandoc')
P <- rprojroot::find_rstudio_root_file
```

###Understanding the ecological drivers of avian influenze virus infection in wildfowl: a continental-scale study across Africa


This detailed and well-sourced paper `r citep('10.1098/rspb.2011.1417')` uses data from tropical wetlands in 15 African countries to evaluate potential ecological drivers of avian influenza prevalence in wild birds. There is no consideration of the interface between wild-bird and domestic-bird, but these results might still be relevant. Their stepwise GLMM model-fitting isolated three factors:  

- Taxonomic group (*Anas* species compared to other species)

- Wildfowl density at the community level, not species level

- Timing of sampling relative to arrival of Eurasian wildfowl, not connected to wet/dry season

The authors discuss several interpretations based on what they found to be significant:

__(1)__ The relative importance of the *Anas*/non-*Anas* dichotomy over the foraging or migratory species behavior of wildfowl suggests that intrinsic biological differences between species may be more important for their receptivity to avian influenza infection. 

__(2)__ The importance of community-level density suggests interspecies mixing is more important than density of individuals; however, when the authors investigated the relationship between the proportion of Eurasian wildfowl to influenza prevalence, they found no correlation -- they conclude that bird origin may not matter as much as the increase in density during these migration times. 

__(3)__ The association between local wildfowl density and influenza infection coupled with the lack of an association between any climatic variables and influenza infection suggests that direct inter-individual transmission is more important for tropical regions than indirect transmission via a persisting environmental reservoir. This contrasts with previous findings for temperate African regions, where indirect transmission is considered more impactful than density.

__(4)__ The lack of (climatic) seasonality as a factor in tropical regions may be because there are extended breeding seasons that produce a gradual increase in susceptible juveniles rather than a spike in births.

```{r, include=FALSE}
write.bibtex(file = P("notebook/references.bib"))
```

## References

<div id="refs"></div>
