---
output: html_document
bibliography: references.bib
csl: elsevier-with-titles.csl
link-citations: true
---

```{r setup, include=FALSE}
library(knitcitations)
cite_options(citation_format = 'pandoc')
P <- rprojroot::find_rstudio_root_file
```

###Assessing the Potential Impact of Avian Influenza on Poultry in West Africa: A Spatial Equilibrium Analysis `r citep('10.1111/j.1477-9552.2007.00099.x')`

This paper uses a spatial equilibrium model to quantify the economic losses that would result from avian influenza outbreaks in West Africa, specifically modeling Nigeria in detail. In the most serious outbreak scenario, where avian influenza is spread widely along two migratory bird pathways in Nigeria and there is local transmission of the disease, the model predicts that the country's chicken production would fall by 21.1% at a cost of around $250 million USD in lost revenue. 

- The authors used FAO *Gridded Livestock of the World* spatial dataset to define poultry distribution. When they compared chicken density and population density, they found that < 4% of West African population lives in high density chicken areas -- most poultry in this region are raised in small backyard groups (< 50 chickens).

- The authors used wild bird flyways as proxy for high risk of avian influenza (using different buffer sizes), assuming that migratory birds transmit the virus to poultry. There was no transmission parameter, just risk buffers.  

- The authors' model only considers poultry-to-poultry and wildfowl-to-poultry transmission, no other human or livestock pathways. 

- This is an economic model which considers supply, demand, and income -- but not import/export of chickens. The demand parameter is "shocked" by the avian influenza outbreak, incorporating the psychological effect an outbreak would have on demand and the chicken economy.

###Modeling highly pathogenic avian influenza transmission in wild birds and poultry in West Bengal, India `r citep('10.1038/srep02175')`

The authors of this paper used a stochastic continuous time model to simulate introduction of H5N1 to poultry communities in West Bengal, India. Additionally, they performed a spatial cluster analysis of outbreaks, calculated R~0~ using three methods, and performed a logistic regression to identify significant predictors of H5N1 outbreaks.

- SIC-SIRS model for change in disease status of a poultry flock (Susceptible - Infectious - Culled) and wild birds (SIRS), unit of interest is the village, with an infectious period of 7 days.  

- Every village (n = 37,945) considered to have both poultry flock and wild birds; once a flock in a village is infected it is assumed entire village is infected and will be culled. 

- Transmission rate between wild birds and poultry communities ($1.00 * 10^{-8}$) was based on previous study of avian influenza dynamics in wild birds `r citep('10.3934/dcdsb.2012.17.2829')`.

- Transmission rate between poultry communities was estimated from SI calculations of observations, similar to a previous paper `r citep('10.1017/S0950268808000885')`, with a mean of 0.127 and exponentially distributed ($\mu = 0.122$).

- R~0~ calculations: 
    - 0.859 (0.326 SD) via  SI equations
    - 1.069 via observed doubling time 
    - 1.034 via simulation results

- No consideration of LPAI in model; only HPAI entering from wild bird groups.


```{r, include=FALSE}
write.bibtex(file = P("notebook/references.bib"))
```

## References

<div id="refs"></div>
