---
title: "Parameters"
bibliography: references.bib
csl: elsevier-with-titles.csl
link-citations: true
output:
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
library(knitcitations)
cite_options(citation_format = 'pandoc')
P <- rprojroot::find_rstudio_root_file
```

###Previous Model Parameters `r citep('10.1371/journal.pone.0080091')`:

Direct Transmission $\beta = 0.004$  

Recovery Rate $\upsilon = (0-0.167)$

Disease Mortality Rate $\alpha = (0 - 0.4)$

Environmental Uptake Rate $\nu = 0.001$

Virion Infectiousness $\phi = 1.96 * 10^{-4}$

Virus Degradation Rate $\eta = 0.14$

Virion Shedding Rate $\sigma = 10^5$

Movement Rate $\omega = 0.30$

###New Model Paramaters

__Poultry to Poultry Direct Transmission $\beta_{PP}$__
Originally adapted from duck study `r citep('10.1128/JVI.78.9.4892-4901.2004')` ~ but unclear how exactly. Existing studies / models deal with village- or flock-level transmission, not individuals. For example, in West Bengal`r citep('10.1038/srep02175')`: $1.00 * 10^{-8}$, originally taken from a study in wild birds `r citep('10.3934/dcdsb.2012.17.2829')`. 

__Wild Bird to Poultry Direct Transmission $\beta_{WP}$__  
Transmission between wild birds and poultry *communities* ($1.00 * 10^{-8}$) was based on previous study of avian influenza dynamics in *wild birds* `r citep('10.3934/dcdsb.2012.17.2829')`

__Poultry to Human Direct Transmission $\beta_{PH}$__  
Fitted parameters from models predicting the seasonality of cumulative cases of HPAI H5N1 `r citep('10.1142/S0218339013400044')`, the estimated range for direct poultry to human transmission was $1.9 * 10^{-11}$ to $2.3 * 10^{-11}$.

__Recovery Rate $\upsilon$__  
Recovery for HPAI infected poultry does not seem to be commonly modeled -- instead either SI models `r citep(list('10.1017/S0950268808000885', '10.1142/S0218339013400044'))` or SIC (Culled) `r citep('10.1038/srep02175')` are used.  

__Disease Mortality Rate $\alpha$__  
Originally given as a range from studies based on ducks / chickens. When re-calculated based on chicken-specific parameter paper`r citep('10.1371/journal.ppat.1000281')`, $\alpha = 0.39$.

__Environmental Uptake Rate $\nu$__  
__Virion Infectiousness $\phi$__  
__Virus Degradation Rate $\eta$__  
__Virion Shedding Rate $\sigma$__  

Environmental transmission not deemed relevant for initial model ~ more relevant for fecal-oral LPAI transmission when no epidemic occurs. In a long-term model could be interesting to consider effects of cleaning. HPAI environmental persistence has been researched under laboratory conditions `r citep('10.1021/es1016153')`. 

__Movement Rate $\omega$__

Original paper uses values from FAO Poultry Review in Egypt`r citep('ftp://ftp.fao.org/docrep/fao/011/ai355e/ai355e00.pdf')`, although exact decision on how to quantify $\omega$ not immediately clear. Once the method is understood, this resource may provide more specific information for Ethiopia: `r citep('http://www.fao.org/docrep/013/al837e/al837e00.pdf')`. Alternatively, FAO data may be relevant here. 


```{r, include=FALSE}
write.bibtex(file = P("notebook/references.bib"))
```

## References

<div id="refs"></div>
