---
output:
  md_document:
    variant: markdown_github
bibliography: notebook/references.bib
csl: notebook/elsevier-with-titles.csl
link-citations: true
---

[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
<a href="https://travis-ci.org/ecohealthalliance/metaflu"> <object data="https://img.shields.io/badge/build-unknown-lightgrey.svg" type="image/png">
    <img src="https://travis-ci.org/ecohealthalliance/metaflu.svg?branch=master" />
  </object></a>
  <a href="https://codecov.io/github/ecohealthalliance/metaflu?branch=master"> <object data="https://img.shields.io/badge/coverage-unknown-lightgrey.svg" type="image/png">
    <img src="https://codecov.io/github/ecohealthalliance/metaflu/coverage.svg?branch=master" />
  </object></a>
[![MIT Licensed - Copyright 2016 EcoHealth Alliance](https://img.shields.io/badge/license-MIT-blue.svg)](https://badges.mit-license.org/)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(knitcitations)
cite_options(citation_format = 'pandoc')
```

# African Sustainable Livestock 2050 Avian Influenza Analysis

This repository contains code and analysis for EcoHealth Alliance's Africa
Sustainable Livestock 2050 project in collaboration with FAO, specifically
for work related to Avian Influenza

This repository is setup as an R package that implements the core model,
which is based on Hosseini et al.'s 2013 metapopulation model 
`r citep("10.1371/journal.pone.0080091")`.  The `notebook/` directory 
contains interim notes and analyses.

```{r bibtex, echo = FALSE, message = FALSE, cache=FALSE, results = 'asis'}
write.bibtex(file="notebook/references.bib")
```

## References

<div id="refs"></div>
