---
title: "Paper Results Reproduction"
author: Cale Basaraba
date: 2016-11-7
output: html_document
bibliography: references.bib
csl: elsevier-with-titles.csl
link-citations: true
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(cache.path='~/.vignette_cache/', echo = FALSE, message=FALSE, warning = FALSE, autodep=TRUE, cache = TRUE)
library(knitcitations)
cite_options(citation_format = 'pandoc')
P <- rprojroot::find_rstudio_root_file
```

Model after `r citep('10.1371/journal.pone.0080091')`.

```{r setup2}
library(metaflu)
library(ggplot2)
library(dplyr)
library(doMC)
library(tidyr)
```


```{r functions}

#function to get duration of epidemic
get_duration <- function(results){
  duration <- results %>%  
    filter(class == "I") %>%
    group_by(sim, time) %>%
    summarize(infectious = sum(population))%>%
    group_by(sim) %>%
    summarize(days_greater = sum(infectious > 0 ))
  duration
}

#function to get cross-sectional infections over time
get_infections <- function(results){
  infections <- results %>%
    filter(class == "I") %>%
    group_by(sim, time) %>%
    summarize(infectious = sum(population)) %>%
    group_by(time) %>%
    summarize(middle = median(infectious), lower = quantile(infectious, probs = c(0.025)), upper = quantile(infectious, probs = c(0.975)))
  infections
}

#function to get cross-sectional susceptible over time
get_susceptibles<- function(results){
  susceptibles <- results %>%
    filter(class == "S") %>%
    group_by(sim, time) %>%
    summarize(susceptibles = sum(population)) %>%
    group_by(time) %>%
    summarize(middle = median(susceptibles), lower = quantile(susceptibles, probs = c(0.025)), upper = quantile(susceptibles, probs = c(0.975)))
  susceptibles
}

# function to get cross-sectional recovered over time
get_recovered<- function(results){
  recovered <- results %>%
    filter(class == "R") %>%
    group_by(sim, time) %>%
    summarize(recovereds = sum(population)) %>%
    group_by(time) %>%
    summarize(middle = median(recovereds), lower = quantile(recovereds, probs = c(0.025)), upper = quantile(recovereds, probs = c(0.975)))
  recovered
}

#function to get epidemic failure rate
get_failure<-function(results){
  check_fails <- function(time, population, patch, class){
    initial<- which(time == 1 & class == "I" & population > 0)  
    all<-which(class == "I" & population > 0)
    identical(patch[initial],unique(patch[all]))
  }
  failures <- results %>%
    group_by(sim) %>%
    summarize(failed = check_fails(time, population, patch, class))
  failures
}

proportion_failed <- function(failure_results){
  sum(failure_results$failed)/length(failure_results$failed)
}

#function to get total number of infections
get_tot_infections <- function(results){
  tot_infections <- results %>%
    filter(class == "S") %>%
    group_by(sim, time) %>%
    summarize(tots = sum(population)) %>%
    group_by(sim) %>%
    summarize(total_i = max(tots) - min(tots))
}


```



```{r figure2-recreate, cache = TRUE, eval = FALSE}
set.seed(17)

fig2 <- function(farm_size, farm_number){
  initial_cond <- matrix(c(farm_size, 0, 0, 0), nrow=farm_number, ncol=4, byrow=TRUE)
  infected_patches <- sample(seq_len(nrow(initial_cond)), 2)
  initial_cond[infected_patches, 2] <- 1
  initial_cond[infected_patches, 1] <- initial_cond[infected_patches, 1] - 1

  fig2parms = list(
    beta = 0.004,   #contact rate for direct transmission
    gamma = 0.167,  #recovery rate
    mu = 0,         #base mortality rate
    alpha = 0.1111,      #disease mortality rate
    phi = 0,  #infectiousness of environmental virions
    eta = 0,     #degradation rate of environmental virions
    nu =  0.00,    #uptake rate of environmental virion
    sigma = 0,      #virion shedding rate
    omega = 0.03,   #movement rate
    rho = 0,        #contact  nonlinearity 0=dens-dependent, 1=freq-dependent
    lambda = 0,     #force of infection from external sources
    chi = matrix(c(1,0,0,1), nrow = 2, ncol = 2),
    network_type = "smallworld",
    network_parms = list(dim = 1, size = farm_number, nei = 2.33, p = 0.0596, multiple = FALSE, loops = FALSE),
    stochastic_network = TRUE
    )

  x <- mf_sim(init = initial_cond, parameters = fig2parms, times=0:1000, n_sims = 100)
  x
}

registerDoMC(cores=35)

eighty_by_fourhundred <- fig2(80, 400)

onesixty_by_twohundred <- fig2(160, 200)

twohundred_by_onesixty <- fig2(200, 160)

threetwenty_by_onehundred <- fig2(320, 100)

fivehundred_by_sixtyfour <- fig2(500, 64)

fig2results <- list(eighty_by_fourhundred, onesixty_by_twohundred, twohundred_by_onesixty, threetwenty_by_onehundred, fivehundred_by_sixtyfour)

saveRDS(fig2results, "fig2results.rds")

```



```{r graph-proportions2}

fig2results <- readRDS("fig2results.rds")

proportions <- unlist(lapply(fig2results, function(x) proportion_failed(get_failure(x))))

df1 <- data.frame(id = c(1:5), fails = proportions)

ggplot(data = df1) +
  geom_line(aes(x = id, y = fails)) +
  scale_x_continuous(name = "Chickens:Farm", labels = c("80:400", "160:200","200:160", "320:100","500:64")) + 
  scale_y_continuous(name = "Proportion of Epidemic Failures")
  theme_bw()
```
  
  
```{r graph-duration2}

duration_list <- lapply(fig2results, get_duration)
lower <- sapply(duration_list,function(x) quantile(unlist(x$days_greater), probs = c(0.025)))
upper <- sapply(duration_list, function(x) quantile(unlist(x$days_greater), probs = c(0.975)))
median <- sapply(duration_list, function(x) quantile(unlist(x$days_greater), probs = c(0.5)))

df2 <- data.frame(lower, median, upper)

ggplot(data = df2) +
  geom_line(aes(x = 1:5, y = median)) +
  geom_line(aes(x = 1:5, y = lower), linetype = "dashed") +
  geom_line(aes(x = 1:5, y = upper), linetype = "dashed") + 
  scale_x_continuous(name = "Chickens:Farm", labels = c("80:400", "160:200","200:160", "320:100","500:64")) + 
  scale_y_continuous(name = "Median Duration (days)") +
  theme_bw()

```




```{r graph-abundance2}

abundance_list <- lapply(fig2results, get_tot_infections)

lower <- sapply(abundance_list,function(x) quantile(unlist(x$total_i), probs = c(0.025)))
upper <- sapply(abundance_list, function(x) quantile(unlist(x$total_i), probs = c(0.975)))
median <- sapply(abundance_list, function(x) quantile(unlist(x$total_i), probs = c(0.5)))

df3 <- data.frame(lower, median, upper)

ggplot(data = df3) +
  geom_line(aes(x = 1:5, y = median)) +
  geom_line(aes(x = 1:5, y = lower), linetype = "dashed") +
  geom_line(aes(x = 1:5, y = upper), linetype = "dashed") + 
  scale_x_continuous(name = "Chickens:Farm", labels = c("80:400", "160:200","200:160", "320:100","500:64")) + 
  scale_y_continuous(name = "Total Number of Infections") +
  theme_bw()


```

The above graph shows that 



```{r graph-sir2, warning = FALSE}
infectious_list <- lapply(fig2results, get_infections)
recovered_list <- lapply(fig2results, get_recovered)
susceptible_list <- lapply(fig2results, get_susceptibles)

s160 <- susceptible_list[[2]]
i160 <- infectious_list[[2]]
r160 <- recovered_list[[2]]

ggplot(data = s160) +
  geom_line(aes(x = time, y = middle), color = "blue") +
  geom_line(data = i160, aes(x = time, y = middle), color = "red") +
  geom_line(data = r160, aes(x = time, y = middle), color = "green") +
  scale_x_continuous(name = "Time (days)", limits = c(0, 500)) + 
  scale_y_continuous(name = "Median Number of Chickens (100 simulations)") 

```

The above graph is an example median SIR graph for a small-world network of two hundred farms with one hundred and sixty chickens on each farm. Susceptible = blue, infected = red, recovered = green.

Median values across simulations aren't accurately representing what's happening here, as the infectious line shows. 




```{r, include=FALSE}
write.bibtex(file="references.bib")
```

## References

<div id="refs"></div>
