---
output:
  html_document:
    keep_md: TRUE
bibliography: references.bib
csl: elsevier-with-titles.csl
link-citations: true
---

```{r setup, include=FALSE}
library(knitcitations)
cite_options(citation_format = 'pandoc')
P <- rprojroot::find_rstudio_root_file
```
#Ecology of Wild Birds
###Victims and vectors: highly pathogenic avian influenza H5N1 and the ecology of wild birds

This review paper `r citep('10.3184/175815510X12737339356701')` examines the ecology and role of wild birds in transmitting HPAI (highly pathogenic avian influenza) to poultry, and ultimately, to humans. Important points include:

- Wild birds serve as the natural reservoir for LPAI (low pathogenic avian influenza) and the authors seem to favor theory that they are primarily shuttles of LPAI (not HPAI) between poultry populations. The authors cite the lack of wild bird HPAI samples and the reduced ability to migrate or function while infected with HPAI to support this theory. Ducks are a notable exception to this, as they can have asymptomatic HPAI infections. 

- The authors argue that HPAI is more likely to occur in systems where wild birds and poultry interact, and cite evidence for HPAI development primarily occurring within poultry hosts for biological reasons.

- Wet markets, rice paddies, and anywhere where backyard flocks commingle with wild birds are identified as areas for HPAI development.

- While LPAI transmission in wild birds is fecal-oral, the authors theorize that HPAI strains adapt toward respiratory transmission due to selection for this trait in closely-kept, confined poultry populations. 

- Cold temperatures help certain influenza subtypes survive in the environment, below average rainfall also correlated with HPAI outbreak in waterbirds. 

- Great tables on key wild bird events, pathologic H5N1 exposure results in bird species, and healthy wild birds found infected with HPAI. 

###Ecology of avian influenza viruses in a changing world
This paper discusses both the ecology of avian influenza viruses and how anthropogenic changes may affect future virus transmission `r citep('10.1111/j.1749-6632.2010.05451.x')`. Some novel points include:

- The authors suggest that shorebirds are likely responsible for distant geographical spread of HPAI. They point out that despite some literature regarding reduced migration ability for infected birds, the evidence is not definitive; additionally, infections could hop between birds in the same flock as the group migrates or be asymptomatic at first, allowing shorebirds to travel large distances. Even given much lower prevalence in these birds, only a few infected cases would allow intercontinental spread.  

- Decreasing wetlands may cause more aggregation of waterfowl and increased contact rate between wild birds and poultry.

- Climate changes in water temperature, higher salinity, and acidic conditions may decrease the survival of influenza virus in the environment.

- Reduced genetic diversity of poultry may make them more susceptible.

- The transit of vaccine-protected poultry may occasionally help spread disease as some vaccine-protected animals continue to shed virus.

###Understanding the ecological drivers of avian influenze virus infection in wildfowl: a continental-scale study across Africa


This well-sourced paper `r citep('10.1098/rspb.2011.1417')` uses data from tropical wetlands in 15 African countries to evaluate potential ecological drivers of avian influenza prevalence in wild birds. There is no consideration of the interface between wild-bird and domestic-bird, but these results might still be relevant. Their stepwise GLMM model-fitting isolated three factors:  

- Taxonomic group (*Anas* species compared to other species)

- Wildfowl density at the community level, not species level

- Timing of sampling relative to arrival of Eurasian wildfowl, not connected to wet/dry season

The authors discuss several interpretations based on what they found to be significant:

__(1)__ The relative importance of the *Anas*/non-*Anas* dichotomy over the foraging or migratory species behavior of wildfowl suggests that intrinsic biological differences between species may be more important for their receptivity to avian influenza infection. 

__(2)__ The importance of community-level density suggests interspecies mixing is more important than density of individuals; however, when the authors investigated the relationship between the proportion of Eurasian wildfowl to influenza prevalence, they found no correlation -- they conclude that bird origin may not matter as much as the increase in density during these migration times. 

__(3)__ The association between local wildfowl density and influenza infection coupled with the lack of an association between any climatic variables and influenza infection suggests that direct inter-individual transmission is more important for tropical regions than indirect transmission via a persisting environmental reservoir. This contrasts with previous findings for temperate African regions, where indirect transmission is considered more impactful than density.

__(4)__ The lack of (climatic) seasonality as a factor in tropical regions may be because there are extended breeding seasons that produce a gradual increase in susceptible juveniles rather than a spike in births.


#Conclusion

The current consensus seems to be that wild birds are the reservoir for LPAI, but it is unclear the degree to which they are responsible for the movement of HPAI from one poultry community to another -- some argue HPAI is most often spread via infected poultry trade and movement. It is possible that wild birds infect poultry with LPAI, which then is converted to HPAI within poultry communities as they infect and re-infect one another. In either scenario, contact with wild birds is the catalyst and therefore remains an important risk factor for initial poultry infection. 


```{r, include=FALSE}
write.bibtex(file = P("vignettes/references.bib"))
```

## References

<div id="refs"></div>
