---
title: "Basic Growth Scenario"
author: "Cale Basaraba"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=10, fig.height=6, warning = FALSE)
devtools::load_all()
library(metaflu)
library(ggplot2)
library(dplyr)
library(doMC)
library(tidyr)
library(purrr)
library(gridExtra)
set.seed(123)
```

```{r patch-functions}

basic_patches <- function(farm_size, farm_number){
  initial_cond <- cbind(rpois(farm_number, farm_size), matrix(0, ncol = 4, nrow = farm_number))
  return(initial_cond)
}

grow_patches <- function(basic_cond){
  s_num <- round(nrow(basic_cond)*1/11)
  new_cond <- basic_cond
  grow_rows <- sample.int(nrow(new_cond), s_num)
  new_cond[grow_rows,1] <- new_cond[grow_rows,1]*10
  return(new_cond)
}

```

### Introduction
This is a bare bones run of the metaflu simulator. 

- ~15 Chickens (Poisson distribution with 15 as mean)

- 100 Farms

- One initial seeded infection


```{r network-setup}
#Set number of farms and ~ number of chickens per farm
farm_number <- 100
farm_size <- 15

#Set up patch size
initial_cond <- basic_patches(farm_size, farm_number)
  infected_patches <- sample(seq_len(nrow(initial_cond)), 1)
  initial_cond[infected_patches, 2] <- 1
  initial_cond[infected_patches, 1] <- initial_cond[infected_patches, 1] - 1

```

#### Parameters

$\beta$ = 0.004 -- Contact rate

$\gamma$ = 0.167 -- Recovery rate

$\mu$ = 0.0 -- Base mortality rate

$\alpha$ = 0.4 -- Disease mortality rate

$\phi$ = 0.0 -- Infectiousness of environmental virions

$\eta$ = 0.0 -- Degradation of environmental virions

$\nu$ = 0.0 -- Update rate of environmental virions

$\sigma$ = 0.0 -- Virion shedding rate

$\omega$ = 0.03 -- Inter-patch movement rate

$\rho$ = 0 -- Contact non-linearity parameter

$\lambda$ = 0.0 -- External force of infection

$\tau$-crit = 0.0 -- Critical surveillance time

$I$-crit = 0.0 -- Threshold for reporting

$\pi$-report = 0.0 -- Reporting probability

$\pi$-detect = 0.0 -- Detection probability

$t_c$-detect = 1  -- Mean time (days) to culling after detection

```{r parameter-setup}
  parms = list(
    beta = 0.004,   #contact rate for direct transmission
    gamma = 0.167,  #recovery rate
    mu = 0,         #base mortality rate
    alpha = 0.4,      #disease mortality rate
    phi = 0,  #infectiousness of environmental virions
    eta = 0,     #degradation rate of environmental virions
    nu =  0.00,    #uptake rate of environmental virion
    sigma = 0,      #virion shedding rate
    omega = 0.03,   #movement rate
    rho = 0,        #contact  nonlinearity 0=dens-dependent, 1=freq-dependent
    lambda = 0,     #force of infection from external sources
    tau_crit = 0,   #critical suveillance time
    I_crit = 0,     #threshold for reporting
    pi_report = 0, #reporting probability
    pi_detect = 0, #detection probability
    cull_time = 1, #culling time
    network_type = "smallworld",
    network_parms = list(dim = 1, size = farm_number, nei = 2.33, p = 0.0596, multiple = FALSE, loops = FALSE),
    stochastic_network = TRUE
    )
```

### Initial Run

```{r basic-run}
basic_results <- mf_sim(init = initial_cond, parameters = parms, times=0:365, n_sims = 1000)

durations <- get_duration_array(basic_results)

s <- get_susceptibles_array(basic_results)
i <- get_infectious_array(basic_results)
r <- get_recovered_array(basic_results)

spread_failures <- get_failure_array(basic_results)

tot_i <- get_tot_infections_array(basic_results)

rm(basic_results)
  
```

```{r plots-basic, message = FALSE}
l <- length(tot_i$total_i)

sir <- ggplot() +
  geom_line(data = s, aes(x = time, y = median, color = "S")) +
  geom_line(data = i, aes(x = time, y = median, color = "I")) +
  geom_line(data = r, aes(x = time, y = median, color = "R")) +
  geom_ribbon(data = s, aes(x = time, ymin = lower, ymax = upper), fill = "blue", alpha = 0.2) +
  geom_ribbon(data = i, aes(x = time, ymin = lower, ymax = upper), fill = "red", alpha = 0.2) +
  geom_ribbon(data = r, aes(x = time, ymin = lower, ymax = upper), fill = "green", alpha = 0.2) +
  labs(title = "S-I-R", x = "Time", y = paste0("Median: ", l, " Sims")) +
  scale_colour_manual(name = "Compartment", values=c(S = "blue", I = "red", R = "green")) +
  scale_x_continuous(limits = c(0, max(durations$duration)+10)) +
  theme_minimal()

total_i <- ggplot() +
  geom_histogram(data = tot_i, aes(x = total_i), bins = 100) +
  labs(title = paste0("Total Infections in ", l," Simulations"), x = "Total Infections", y = "Number of Simulations") +
  theme_minimal()

failures <- ggplot() +
  geom_bar(data = spread_failures, aes(x = 1, fill = failed), width = 0.1) +
  theme_minimal() +
  labs(title = "Failure to Spread", y = "Number of Simulations") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  scale_fill_manual(name = "Failed to Spread", values = c("black", "grey")) 

duration <- ggplot() +
  geom_histogram(data = durations, aes(x = durations$duration), binwidth = 0.5) +
  theme_minimal() +
  labs(title = paste0("Duration of Epidemic in ", l, " Simulations"), x = "Duration in Days", y = "Number of Simulations")

lay <- rbind(c(1,1,1,4),
             c(1,1,1,4),
             c(2,2,3,3))

grid.arrange(sir,total_i,duration,failures, layout_matrix = lay)
```

### Growth Run

The only change made to this run is that approximately 11% of the original farms grow to ten-times their original size (~15 to ~150), resulting in twice as many chickens in the entire network. For this experiment, the growing farms are chosen at random.

```{r growth-run}
#Grow a random number of patches to get double total chicken population
grown_cond <- grow_patches(initial_cond)

grown_results <- mf_sim(init = grown_cond, parameters = parms, times = 0:365, n_sims = 1000)

g_durations <- get_duration_array(grown_results)

g_s <- get_susceptibles_array(grown_results)
g_i <- get_infectious_array(grown_results)
g_r <- get_recovered_array(grown_results)

g_spread_failures <- get_failure_array(grown_results)

g_tot_i <- get_tot_infections_array(grown_results)

rm(grown_results)

```

```{r plots-growth, message = FALSE}
l <- length(g_tot_i$total_i)

g_sir <- ggplot() +
  geom_line(data = g_s, aes(x = time, y = median, color = "S")) +
  geom_line(data = g_i, aes(x = time, y = median, color = "I")) +
  geom_line(data = g_r, aes(x = time, y = median, color = "R")) +
  geom_ribbon(data = g_s, aes(x = time, ymin = lower, ymax = upper), fill = "blue", alpha = 0.2) +
  geom_ribbon(data = g_i, aes(x = time, ymin = lower, ymax = upper), fill = "red", alpha = 0.2) +
  geom_ribbon(data = g_r, aes(x = time, ymin = lower, ymax = upper), fill = "green", alpha = 0.2) +
  labs(title = "S-I-R", x = "Time", y = paste0("Median: ", l, " Sims")) +
  scale_colour_manual(name = "Compartment", values=c(S = "blue", I = "red", R = "green")) +
  scale_x_continuous(limits = c(0, max(g_durations$duration)+10)) +
  theme_minimal()

g_total_i <- ggplot() +
  geom_histogram(data = g_tot_i, aes(x = total_i), bins = 100) +
  labs(title = paste0("Total Infections in ", l," Simulations"), x = "Total Infections", y = "Number of Simulations") +
  theme_minimal()

g_failures <- ggplot() +
  geom_bar(data = g_spread_failures, aes(x = 1, fill = failed), width = 0.1) +
  theme_minimal() +
  labs(title = "Failure to Spread", y = "Number of Simulations") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  scale_fill_manual(name = "Failed to Spread", values = c("black", "grey")) 

g_duration <- ggplot() +
  geom_histogram(data = g_durations, aes(x = g_durations$duration), binwidth = 0.5) +
  theme_minimal() +
  labs(title = paste0("Duration of Epidemic in ", l, " Simulations"), x = "Duration in Days", y = "Number of Simulations")

lay <- rbind(c(1,1,1,4),
             c(1,1,1,4),
             c(2,2,3,3))

grid.arrange(g_sir,g_total_i,g_duration,g_failures, layout_matrix = lay)


```


