---
title: "External Force of Infection"
author: "Cale Basaraba"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=10, fig.height=6, warning = FALSE)
devtools::load_all()
library(metaflu)
library(ggplot2)
library(dplyr)
library(doMC)
library(tidyr)
library(purrr)
library(gridExtra)
library(grid)
set.seed(123)
```

```{r patch-functions}

basic_patches <- function(farm_size, farm_number){
  initial_cond <- cbind(rpois(farm_number, farm_size), matrix(0, ncol = 3, nrow = farm_number))
  return(initial_cond)
}

grow_patches <- function(basic_cond){
  s_num <- round(nrow(basic_cond)*1/9)
  new_cond <- basic_cond
  grow_rows <- sample.int(nrow(new_cond), s_num)
  new_cond[grow_rows,1] <- new_cond[grow_rows,1]*10
  return(new_cond)
}

```

### Introduction

This goal of this document is to determine an appropriate $\lambda$-value that represents a constant external force of infection. 

#### Expected number of epidemics

The following set of simulations includes the following scenario options:

- Non-linear Contact Rate ($\rho, \beta$)

- Clustered Intensification in Growth Scenario

- Paired Networks: Initial / Growth Scenarios have same set of stochastic networks

- Poisson distribution with 40 as mean number of chickens per farm

- 100 Farms

- One initial seeded infection


```{r expected-setup}
#Set number of farms and ~ number of chickens per farm
farm_number <- 100
farm_size <- 40

#Set up patch size
initial_cond <- basic_patches(farm_size, farm_number)
  infected_patches <- sample(seq_len(nrow(initial_cond)), 1)
  initial_cond[infected_patches, 2] <- 1
  initial_cond[infected_patches, 1] <- initial_cond[infected_patches, 1] - 1

```

#### Expected Model Parameters

$\beta$ = 1.4456 -- Contact rate implemented alongside $\rho$

$\gamma$ = 0.167 -- Recovery rate

$\mu$ = 0.0 -- Base mortality rate

$\alpha$ = 0.4 -- Disease mortality rate

$\phi$ = 0.0 -- Infectiousness of environmental virions

$\eta$ = 0.0 -- Degradation of environmental virions

$\nu$ = 0.0 -- Update rate of environmental virions

$\sigma$ = 0.0 -- Virion shedding rate

$\omega$ = 0.03 -- Inter-patch movement rate

$\rho$ = 0.8526 -- Contact non-linearity parameter

$\lambda$ = 0.0 -- External force of infection

$\tau$-crit = 0.0 -- Critical surveillance time

$I$-crit = 0.0 -- Threshold for reporting

$\pi$-report = 0.0 -- Reporting probability

$\pi$-detect = 0.0 -- Detection probability

```{r expected-parameters}
  exp_parm = list(
    beta = 1.44456,   #contact rate for direct transmission
    gamma = 0.167,  #recovery rate
    mu = 0,         #base mortality rate
    alpha = 0.4,      #disease mortality rate
    phi = 0,  #infectiousness of environmental virions
    eta = 0,     #degradation rate of environmental virions
    nu =  0.00,    #uptake rate of environmental virion
    sigma = 0,      #virion shedding rate
    omega = 0.03,   #movement rate
    rho = 0.85256,        #contact  nonlinearity 0=dens-dependent, 1=freq-dependent
    lambda = 0,     #force of infection from external sources
    tau_crit = 0,   #critical suveillance time
    I_crit = 0,     #threshold for reporting
    pi_report = 0, #reporting probability
    pi_detect = 0, #detection probability
    network_type = "smallworld",
    network_parms = list(dim = 1, size = farm_number, nei = 2.33, p = 0.0596, multiple = FALSE, loops = FALSE),
    stochastic_network = TRUE
    )
```


```{r expected-run}
basic_results <- mf_sim(init = initial_cond, parameters = exp_parm, times=1:365, n_sims = 1000)

durations <- get_duration_array(basic_results)

epi_start <- get_epi_begin(basic_results)

s.df <- truncate_data(epi_start, get_all_sims("S", basic_results))
i.df <- truncate_data(epi_start, get_all_sims("I", basic_results))
r.df <- truncate_data(epi_start, get_all_sims("R", basic_results))

spread_failures <- get_failure_array(basic_results)

tot_i <- get_tot_infections_array(basic_results)

nets <- attr(basic_results, "network")

rm(basic_results)
  
```

```{r plots-basic, message = FALSE}
l <- length(tot_i$total_i)


sir.df <- ggplot() +
  geom_line(data = s.df, aes(x = new_time, y = pop, group = sim), alpha = 0.05, color = "blue") +
  geom_line(data = i.df, aes(x = new_time, y = pop, group = sim), alpha = 0.05, color = "red") +
  geom_line(data = r.df, aes(x = new_time, y = pop, group = sim), alpha = 0.05, color = "green") +
  labs(title = "S-I-R", x = "Time", y = "Population") +
  scale_colour_manual(name = "Compartment", values=c(S = "blue", I = "red", R = "green")) +
  theme_minimal()

#s.df2 <- filter(s.df, sim == 2)
#i.df2 <- filter(i.df, sim == 2)
#r.df2 <- filter(r.df, sim == 2)


#sir.df2 <- ggplot() +
#  geom_line(data = s.df2, aes(x = new_time, y = pop, group = sim), alpha = 0.5, color = "blue") +
#  geom_line(data = i.df2, aes(x = new_time, y = pop, group = sim), alpha = 0.5, color = "red") +
#  geom_line(data = r.df2, aes(x = new_time, y = pop, group = sim), alpha = 0.5, color = "green") +
#  labs(title = "S-I-R", x = "Time", y = "Population") +
#  scale_colour_manual(name = "Compartment", values=c(S = "blue", I = "red", R = "green")) +
#  theme_minimal()

total_i <- ggplot() +
  geom_histogram(data = tot_i, aes(x = total_i), bins = 100) +
  labs(title = paste0("Total Infections in ", l," Simulations"), x = "Total Infections", y = "Number of Simulations") +
  theme_minimal()

failures <- ggplot() +
  geom_bar(data = spread_failures, aes(x = 1, fill = failed), width = 0.1) +
  theme_minimal() +
  labs(title = "Failure to Spread", y = "Number of Simulations") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  scale_fill_manual(name = "Failed to Spread", values = c("black", "grey")) 

failures <- textGrob("Failures")

duration <- ggplot() +
  geom_histogram(aes(x = durations), binwidth = 0.5) +
  theme_minimal() +
  labs(title = paste0("Duration of Epidemic in ", l, " Simulations"), x = "Duration in Days", y = "Number of Simulations")

lay <- rbind(c(1,1,1,4),
             c(1,1,1,4),
             c(2,2,3,3))

grid.arrange(sir.df,total_i,duration,failures, layout_matrix = lay)

```

#### Calculating Lambda

Based on these seeded simulations, the expected proportion of epidemics (simulations which spread beyond one farm), is `r 1-sum(spread_failures$failed)/length(spread_failures$failed)`.

In order to match this expected proportion over a one year period, $\lambda = 0.684 * \frac{1}{100}\ * \frac{1}{40} * \frac{1}{365} = 4.684932 * 10^{-7}$

The following run simulates the same scenario with this external force of infection:


### External Infection Run

```{r network-setup}
#Set number of farms and ~ number of chickens per farm
farm_number <- 100
farm_size <- 40

#Set up patch size
initial_cond <- basic_patches(farm_size, farm_number)

```

#### Parameters

$\beta$ = 1.4456 -- Contact rate implemented alongside $\rho$

$\gamma$ = 0.167 -- Recovery rate

$\mu$ = 0.0 -- Base mortality rate

$\alpha$ = 0.4 -- Disease mortality rate

$\phi$ = 0.0 -- Infectiousness of environmental virions

$\eta$ = 0.0 -- Degradation of environmental virions

$\nu$ = 0.0 -- Update rate of environmental virions

$\sigma$ = 0.0 -- Virion shedding rate

$\omega$ = 0.03 -- Inter-patch movement rate

$\rho$ = 0.8526 -- Contact non-linearity parameter

$\lambda$ = 4.684932e-07 -- External force of infection

$\tau$-crit = 0.0 -- Critical surveillance time

$I$-crit = 0.0 -- Threshold for reporting

$\pi$-report = 0.0 -- Reporting probability

$\pi$-detect = 0.0 -- Detection probability

```{r parameter-setup}
  parms = list(
    beta = 1.44456,   #contact rate for direct transmission
    gamma = 0.167,  #recovery rate
    mu = 0,         #base mortality rate
    alpha = 0.4,      #disease mortality rate
    phi = 0,  #infectiousness of environmental virions
    eta = 0,     #degradation rate of environmental virions
    nu =  0.00,    #uptake rate of environmental virion
    sigma = 0,      #virion shedding rate
    omega = 0.03,   #movement rate
    rho = 0.85256,        #contact  nonlinearity 0=dens-dependent, 1=freq-dependent
    lambda = 4.684932e-07,     #force of infection from external sources
    tau_crit = 0,   #critical suveillance time
    I_crit = 0,     #threshold for reporting
    pi_report = 0, #reporting probability
    pi_detect = 0, #detection probability
    network_type = "smallworld",
    network_parms = list(dim = 1, size = farm_number, nei = 2.33, p = 0.0596, multiple = FALSE, loops = FALSE),
    stochastic_network = TRUE
    )
```

### Initial Run

```{r basic-run}

basic_results <- mf_sim(init = initial_cond, parameters = parms, times=1:365, n_sims = 1000)

epi_start <- get_epi_begin(basic_results)

s.df <- truncate_data(epi_start, get_all_sims("S", basic_results))
i.df <- truncate_data(epi_start, get_all_sims("I", basic_results))
r.df <- truncate_data(epi_start, get_all_sims("R", basic_results))

durations <- get_duration_lambda(basic_results)

spread_failures <- get_failure_array(basic_results)

tot_i <- get_tot_infections_array(basic_results)

#rm(basic_results)
  

```

```{r plots-basic, message = FALSE}
l <- length(tot_i$total_i)


sir.df <- ggplot() +
  geom_line(data = s.df, aes(x = new_time, y = pop, group = sim), alpha = 0.05, color = "blue") +
  geom_line(data = i.df, aes(x = new_time, y = pop, group = sim), alpha = 0.05, color = "red") +
  geom_line(data = r.df, aes(x = new_time, y = pop, group = sim), alpha = 0.05, color = "green") +
  labs(title = "S-I-R", x = "Time", y = "Population") +
  scale_colour_manual(name = "Compartment", values=c(S = "blue", I = "red", R = "green")) +
  theme_minimal()

#s.df2 <- filter(s.df, sim == 2)
#i.df2 <- filter(i.df, sim == 2)
#r.df2 <- filter(r.df, sim == 2)


#sir.df2 <- ggplot() +
#  geom_line(data = s.df2, aes(x = new_time, y = pop, group = sim), alpha = 0.5, color = "blue") +
#  geom_line(data = i.df2, aes(x = new_time, y = pop, group = sim), alpha = 0.5, color = "red") +
#  geom_line(data = r.df2, aes(x = new_time, y = pop, group = sim), alpha = 0.5, color = "green") +
#  labs(title = "S-I-R", x = "Time", y = "Population") +
#  scale_colour_manual(name = "Compartment", values=c(S = "blue", I = "red", R = "green")) +
#  theme_minimal()

total_i <- ggplot() +
  geom_histogram(data = tot_i, aes(x = total_i), bins = 100) +
  labs(title = paste0("Total Infections in ", l," Simulations"), x = "Total Infections", y = "Number of Simulations") +
  theme_minimal()

failures <- ggplot() +
  geom_bar(data = spread_failures, aes(x = 1, fill = failed), width = 0.1) +
  theme_minimal() +
  labs(title = "Failure to Spread", y = "Number of Simulations") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  scale_fill_manual(name = "Failed to Spread", values = c("black", "grey")) 

failures <- textGrob("Failures")

duration <- ggplot() +
  geom_histogram(aes(x = durations), binwidth = 0.5) +
  theme_minimal() +
  labs(title = paste0("Duration of Epidemic in ", l, " Simulations"), x = "Duration in Days", y = "Number of Simulations")

lay <- rbind(c(1,1,1,4),
             c(1,1,1,4),
             c(2,2,3,3))

grid.arrange(sir.df,total_i,duration,failures, layout_matrix = lay)

```
