---
title: "Delayed Culling Check"
author: "Cale Basaraba"
date: "4/6/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=10, fig.height=6, warning = FALSE)
library(metaflu)
library(ggplot2)
library(dplyr)
library(doMC)
library(tidyr)
library(purrr)
library(gridExtra)
library(abind)
library(grid)
set.seed(123)
registerDoMC(cores = detectCores())
```


```{r param-setup}
farm_number <- 100
farm_size <- 40

parms = list(
  beta = 1.44456,   #contact rate for direct transmission
  gamma = 0.167,  #recovery rate
  mu = 0,         #base mortality rate
  alpha = 0.4,      #disease mortality rate
  phi = 0,  #infectiousness of environmental virions
  eta = 0,     #degradation rate of environmental virions
  nu =  0.00,    #uptake rate of environmental virion
  sigma = 0,      #virion shedding rate
  omega = 0.03,   #movement rate
  rho = 0.85256,        #contact  nonlinearity 0=dens-dependent, 1=freq-dependent
  lambda = 0,     #force of infection from external sources
  tau_crit = 1,   #critical suveillance time
  I_crit = 5,     #threshold for reporting
  pi_report = 0.9, #reporting probability
  pi_detect = 0.9, #detection probability
  cull_time = 1, #detection cull time
  network_type = "smallworld",
  network_parms = list(dim = 1, size = farm_number, nei = 2.33, p = 0.0596, multiple = FALSE, loops = FALSE),
  stochastic_network = TRUE
)
```


```{r sims, include = FALSE}
sims <- 100

get_plots <- function(val){
  parms["cull_time"] <- val
  sim_list <- mclapply(seq_len(sims), function(num){
    conditions <- create_initial_condition(basic_patches(farm_size, farm_number))
    return(mf_sim(init = conditions, parameters = parms, times=1:365, n_sims = 1))
  }, mc.cores = detectCores())
  basic_results <- do.call("abind", sim_list)

  # RETURNS
  durations <- get_duration_array(basic_results)
  dur_plot <- ggplot() + geom_histogram(data = durations, aes(x = duration), bins = 100) +
    labs(x = "Duration in Days", y = "Number of Simulations") +
    theme_classic()

  farm_no <- get_number_farms(basic_results)
  farm_plot <- ggplot() + geom_histogram(aes(x = farm_no), bins = 100) +
    labs(x = "Number of Farms Infected", y = "Number of Simulations") +
    theme_classic()

  tot_i <- get_tot_infections_array(basic_results)
  i_plot <- ggplot() + geom_histogram(aes(x = tot_i$total_i), bins = 100) +
    labs(x = "Number of Farms Infected", y = "Number of Simulations") + 
    theme_classic()
  
  tot_culls <- get_number_culls(basic_results)
  cull_plot <- ggplot() + geom_histogram(aes(x = tot_culls), bins = 100) +
        labs(x = "Number of Farms Culled", y = "Number of Simulations") + 
    theme_classic()
  
  lay <- rbind(c(1,1,2,2),
               c(3,3,4,4))

  arranged_plot <- grid.arrange(i_plot, dur_plot, farm_plot, cull_plot, layout_matrix = lay, top = paste("Culling time =",val,"days"))
  gb <- rectGrob(height = 1, width = 1, gp = gpar(lwd = 2, col = "dimgrey", fill = NA))
  gt <- gTree(children = gList(arranged_plot, gb))
  grid.draw(gt)
}
vals <- seq_len(10)


```

###Parameters
The parameters for these runs:

`r farm_size` chickens

`r farm_number` farms

`r sims` simulations per parameter set.

$\beta$ = `r parms["beta"]` -- Contact rate

$\gamma$ = `r parms["gamma"]` -- Recovery rate

$\mu$ = `r parms["mu"]` -- Base mortality rate

$\alpha$ = `r parms["alpha"]` -- Disease mortality rate

$\phi$ = `r parms["phi"]` -- Infectiousness of environmental virions

$\eta$ = `r parms["eta"]` -- Degradation of environmental virions

$\nu$ = `r parms["nu"]` -- Update rate of environmental virions

$\sigma$ = `r parms["sigma"]` -- Virion shedding rate

$\omega$ = `r parms["omega"]` -- Inter-patch movement rate

$\rho$ = `r parms["rho"]` -- Contact non-linearity parameter

$\lambda$ = `r parms["lambda"]` -- External force of infection

$\tau$-crit = `r parms["tau_crit"]` -- Critical surveillance time

$I$-crit = `r parms["I_crit"]` -- Threshold for reporting

$\pi$-report = `r parms["pi_report"]` -- Reporting probability

$\pi$-detect = `r parms["pi_detect"]` -- Detection probability

$t_c$-detect = *1 to 10, by 1  -- Mean time (days) to culling after detection*



```{r plots, include = FALSE, eval = FALSE}
plot_test <- lapply(vals, function(x) get_plots(x))
```


```{r per-day}
cull_time_vals <- seq_len(10)

cull_time_plots <- function(val){
parms["cull_time"] <- val
conditions <- create_initial_condition(basic_patches(farm_size, farm_number))
res1 <- mf_sim(init = conditions, parameters = parms, times=1:365, n_sims = 1000)

count_ones <- apply(res1,4, function(x){
  c <- x["C",,]
  ones <- apply(c, 1, function(y){
    sum(y == 1)
  })
})

ones <- as.vector(count_ones)
only_ones <- ones[which(ones > 0)]

ggplot() + geom_histogram(aes(x = only_ones), binwidth = 0.5) +
  labs(x = "Number of Days from Reported to Culled State", y = "Number of Patches", title = "Time to Culling", subtitle = paste("Mean =", round(mean(only_ones),2), "Parameter =", val)) +
  theme_minimal()
}

```

```{r time_plots}
lapply(cull_time_vals, function(x) cull_time_plots(x))

```